---
title: "Footes_Formulas"
author: "Elijah Phifer"
date: "2023-07-14"
output: html_document
---

# Foote Functions

This information comes from Mike Foote's paper: On the Probability of Ancestors in the Fossil Record.

## Introduction

There are several functions in this script that relate to species preservation probability calculations. This documentation provides an overview of each function.

## `PDT(q, p, R, t)`

This function calculates the probability that the duration of a species is T time units.

- **Parameters:**
  - `q`: The extinction rate of the species.
  - `p`: The origination rate of the species.
  - `R`: The probability of preserving a species at least once within the time it exists.
  - `t`: The duration for which preservation probability is calculated.

- **Returns:**
  The calculated probability value.

## `RTerm(q, p, R, t)`

This function computes the value of BigR, which is a factor that represents represents the probability of preserving a species once.

- **Parameters:**
  - `q`: The extinction rate of the species.
  - `p`: The origination rate of the species.
  - `R`: The probability of preserving a species at least once within the time it exists.
  - `t`: The duration for which preservation probability is calculated.

- **Returns:**
  The value of BigR (the R factor in Foote's Formulas).

## `species_preserved(q, p, R, t)`

This function calculates proportion of species preserved.

- **Parameters:**
  - `q`: The extinction rate of the species.
  - `p`: The origination rate of the species.
  - `R`: The probability of preserving a species at least once within the time it exists.
  - `t`: The duration for which preservation probability is calculated.

- **Returns:**
  The proportion calculation.
  
```{r}
# Example 1: Calculate the probability of species preservation
preservation_sum <- species_preserved(q = 0.1, p = 0.2, R = 0.8, t = 2)
print(preservation_sum)
```

## `PoN(q, p)`

This function calculates the probability that a species with duration T has exactly N direct decendents.

- **Parameters:**
  - `q`: The extinction rate of the species.
  - `p`: The origination rate of the species.

- **Returns:**
  The calculated proportion value.

## `nPP(q, p, R, t, N)`

This function calculates the probability that at least one direct decendent is preserved.

- **Parameters:**
  - `q`: The extinction rate of the species.
  - `p`: The origination rate of the species.
  - `R`: The probability of preserving a species at least once within the time it exists.
  - `t`: The duration for which preservation probability is calculated.
  - `N`: The number of direct decedents.

- **Returns:**
  The calculated probability value.

## `NerSum(q, p, R, t, N)`

This function calculates the sum of the inner sum with in the Pa formula in Foote's paper.

- **Parameters:**
  - `q`: The extinction rate of the species.
  - `p`: The origination rate of the species.
  - `R`: The probability of preserving a species at least once within the time it exists.
  - `t`: The duration for which preservation probability is calculated.
  - `N`: The number of direct decedents.

- **Returns:**
  The sum of the inner sum with in the Pa formula in Foote's paper..

## `decendents_presevered(q, p, R, t, N)`

This function calculates the proportion of species preserved that have at least one direct decedent preserved.

- **Parameters:**
  - `q`: The extinction rate of the species.
  - `p`: The origination rate of the species.
  - `R`: The probability of preserving a species at least once within the time it exists.
  - `t`: The duration for which preservation probability is calculated.
  - `N`: The number of direct decedents.

- **Returns:**
  The proportion of species preserved that have at least one direct decedent preserved.
  
```{r}
# Example 5: Calculate the proportion of species preserved at least once out of N species
descendants_sum <- decendents_presevered(q = 0.1, p = 0.2, R = 0.8, t = 2, N = 100)
print(descendants_sum)
```

---
