---
title: "BF_calulations"
author: "April Wright"
date: "2023-11-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's get some Bayes Factors from our posterior traces.

```{r}
library(tidyverse)
library(RevGadgets)
```

We start off by reading in our trace:

```{r}
files <- c("fbd_jump.log")
trace <- readTrace(path = files, burnin = 0.3)
```


Next, some helper variables. These are the log BF scores that correspond to weak, substantial, and strong BF support.

```{r}
BF <- c(3.2, 10, 100)
p = BF/(1+BF)
```

Next, this takes a bit of doing. We need to know the proportion of the posterior that each model takes up. 

```{r}
big_trace <- rbind(trace[[1]], trace[[2]])
count <- big_trace  %>% 
     group_by(character_indicator) %>% 
     summarize(count = n())
count$count[1]/sum(count)
count$count[2]/sum(count)
```
This proportion is the Bayes Factor. Now, we need to take this number and compare it to a test statistic using the vector of Bayes Factor test statistics (p) above.

On this plot, each model will be a bar. The prior (equal probability) will be in black. The red lines are support (lowest - weak support, the middle - substantial, the top - strong) The green ones are anti-support. Not sure we can see these with how these are calculated.

```{r}

plotTrace(trace = trace,
          vars = c("character_indicator"))[[1]] +
          ylim(0,1) +
          geom_hline(yintercept=p, linetype=c("longdash","dashed","dotted"), color = "black") +
          geom_hline(yintercept=1/5, color = "black") +
     # modify legend location using ggplot2
     theme(legend.position = c(0.40,0.825))

#ggsave(filename = "solocharactermodel.png", width = 7, height = 7)
```
Note that for the FBD, we have 3 models. Thus, a flat prior should be at .33.
```{r}
BF <- c(3.2, 10, 100)
p = (BF/(1+BF))
plotTrace(trace = trace,
          vars = c("fbd_indicator"))[[1]] +
          ylim(0,1) +
          geom_hline(yintercept=0.33, linetype="solid", color = "black") +
          geom_hline(yintercept=p, linetype=c("longdash","dashed","dotted"), color = "red") +
          geom_hline(yintercept=1-p, linetype=c("longdash","dashed","dotted"), color = "green") +
     # modify legend location using ggplot2
     theme(legend.position = c(0.40,0.825))
#ggsave(filename = "solofbdmodel.png", width = 7, height = 7)

```

And the clock:

Note that for the FBD, we have 3 models. Thus, a flat prior should be at .33.You'll notice the ACLN was sampled so infrequently it doesn't show.
```{r}
plotTrace(trace = trace,
          vars = c("clock_indicator"))[[1]] +
          ylim(0,1) +
          geom_hline(yintercept=0.33, linetype="solid", color = "black") +
          geom_hline(yintercept=p, linetype=c("longdash","dashed","dotted"), color = "red") +
          geom_hline(yintercept=1-p, linetype=c("longdash","dashed","dotted"), color = "green") +
     # modify legend location using ggplot2
     theme(legend.position = c(0.40,0.825))
#ggsave(filename = "soloclockmodel.png", width = 7, height = 7)

```

## Treespace plots

First, we load in the trees. 

```{r}

my.trees <- load.multi("../output/rjMCMC_components/fbd/", format = "revbayes", trim = 10)
my.treespace <- makeplot.treespace(my.trees, n.points = 100,
                    fill.color = NA)

```


```{r}
qplot(x, y, data = my.treespace$treespace.points.plot$data, color = chain) +theme_bw()
ggsave("CompoJump.png")
```